---
title: "Sesión 2. Self-organizing maps"
author: "Aprendizaje Máquina II. Máster Ciencia Datos. Universidad de Valencia."
output:
  html_document:
    number_sections: yes
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Iris data preparation

```{r}
library(aweSOM)
data("iris")
full.data <- iris
train.data <- full.data[, c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")]
train.data <- scale(train.data)

cor_matrix<-cor(train.data)
print(cor_matrix)
e<-eigen(cor_matrix)
print(e[1])
```

## Training

```{r}
### RNG Seed (for reproducibility)

set.seed(1465)
### Initialization (PCA grid)
init <- somInit(train.data, 12, 6)
## Train SOM

# alpha = learning_rate (learning rate, a vector of two numbers indicating the amount of change. 
# rlen = the number of times the complete data set will be presented to the network.
# radius = the radius of the neighbourhood, either given as a single number or a vector (start, stop).
# dist.fcts = vector of distance functions to be used for the individual data layer.
iris.som <- kohonen::som(train.data, grid = kohonen::somgrid(12, 6, "rectangular"), 
                         rlen = 100, alpha = c(0.05, 0.01), radius = c(2.65,-2.65), 
                         dist.fcts = "sumofsquares", init = init)
```

## Quality of the SOM

```{r}

num <- somQuality(iris.som, train.data)

print(num$err.quant)
print(num$err.kaski)
print(num$err.topo)
print(num$err.varratio)

```

## Superclass

PAM (k-medians) and Hierarchical clustering. 3 superclasses are selected.

```{r}
superclust_pam <- cluster::pam(iris.som$codes[[1]], 3)
superclasses_pam <- superclust_pam$clustering

superclust_hclust <- hclust(dist(iris.som$codes[[1]]), "complete")
superclasses_hclust <- cutree(superclust_hclust, 3)
```

## Plotting

Observations cloud

```{r}
aweSOMplot(som = iris.som, type = "Cloud", data = full.data, 
           variables = c("Species", "Sepal.Length", "Sepal.Width",  
                         "Petal.Length", "Petal.Width"), 
           superclass = superclasses_pam)
```

Hitmap o population map

```{r}
aweSOMplot(som = iris.som, type = "Hitmap", superclass = superclasses_pam)

```

U-Matrix

```{r}
aweSOMplot(som = iris.som, type = "UMatrix", superclass = superclasses_pam)
```

Smooth U-matrix. Note, however, that the result representation is biased when using hexagonal maps (the smoothing function coerces the grid to a rectangular shape).

```{r}
aweSOMsmoothdist(iris.som)
```

## Plotting numeric variables on the map

Pie plots

```{r}
aweSOMplot(som = iris.som, type = "Circular", data = full.data, 
           variables = c("Sepal.Length", "Sepal.Width",  
                         "Petal.Length", "Petal.Width"), 
           superclass = superclasses_pam)
```

Barplot

```{r}
aweSOMplot(som = iris.som, type = "Barplot", data = full.data, 
           variables = c("Sepal.Length", "Sepal.Width",  
                         "Petal.Length", "Petal.Width"), 
           superclass = superclasses_pam, 
           values = "prototypes")
```

Boxplot

```{r}
aweSOMplot(som = iris.som, type = "Boxplot", data = full.data, 
           variables = c("Sepal.Length", "Sepal.Width",  
                         "Petal.Length", "Petal.Width"),
           superclass = superclasses_pam, 
           scales = "same")
```

Line plot

```{r}
aweSOMplot(som = iris.som, type = "Line", data = full.data, 
           variables = c("Sepal.Length", "Sepal.Width",  
                         "Petal.Length", "Petal.Width"), 
           superclass = superclasses_pam, 
           values = "median")
```

Radar chart

```{r}
aweSOMplot(som = iris.som, type = "Radar", data = full.data, 
           variables = c("Sepal.Length", "Sepal.Width",  
                         "Petal.Length", "Petal.Width"), 
           superclass = superclasses_pam)
```

Color plot

```{r}
aweSOMplot(som = iris.som, type = "Color", data = full.data, 
           variables = "Sepal.Length", superclass = superclasses_pam)
```

## Plotting a categorical variable on the map

Barplot

```{r}
aweSOMplot(som = iris.som, type = "CatBarplot", data = full.data, 
           variables = "Species", superclass = superclasses_pam)
```

Pie plot

```{r}
aweSOMplot(som = iris.som, type = "Pie", data = full.data, variables = "Species", 
           superclass = superclasses_pam)
```

## Choosing the number of superclasses

`aweSOMscreeplot` produces a scree plot, that shows the quality of the clustering (percentage of unexplained variance of the prototypes, lower is better) for varying numbers of superclasses. It supports hierarchical and pam clustering. The rule of thumb is to choose the number of superclasses at the inflection point of this curve.

```{r}
aweSOMscreeplot(som = iris.som, method = "pam", nclass = 3)
```

`aweSOMsilhouette` returns a silhouette plot of the chosen superclasses (hierarchical, pam, or other). The higher the silhouettes, the better.

```{r}
aweSOMsilhouette(iris.som, superclasses_pam)
```

For hierarchical clustering, `aweSOMdendrogram` produces a dendogram, along with the chosen cuts.

```{r}
aweSOMdendrogram(clust = superclust_hclust, nclass = 3)
```
