---
title: "Clasificador basado en reglas Mushroom"
subtitle: "Aprendizaje Máquina II"
author: "Máster Ciencia Datos- UV"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Datos

**Mushroom Data Set**

<https://archive.ics.uci.edu/ml/datasets/mushroom>

Mushroom records drawn from The Audubon Society Field Guide to North
American Mushrooms (1981). G. H. Lincoff (Pres.), New York: Alfred A.
Knopf

```{r}
rm(list=ls())
load(file="mushroom.RData")
```

## Distribución de la clase 

**Comprueba que la distribución de las setas es:**

-- edible: 4208 (51.8%)

-- poisonous: 3916 (48.2%)

-- total: 8124 instances

```{r}
prop.table(table(mushroom$class))

```

## Data Set Information

This data set includes descriptions of hypothetical samples
corresponding to 23 species of gilled mushrooms in the Agaricus and
Lepiota Family. Each species is identified as definitely edible,
definitely poisonous, or of unknown edibility and not recommended. This
latter class was combined with the poisonous one. The Guide clearly
states that there is no simple rule for determining the edibility of a
mushroom; no rule like \`\`leaflets three, let it be'' for Poisonous Oak
and Ivy.

Variables

1\. cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s

2\. cap-surface: fibrous=f,grooves=g,scaly=y,smooth=s

3\. cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r,
pink=p,purple=u,red=e,white=w,yellow=y

4\. bruises?: bruises=t,no=f

5\. odor: almond=a, anise=l, creosote=c, fishy=y, foul=f, musty=m,
none=n, pungent=p, spicy=s

6\. gill-attachment: attached=a,descending=d,free=f,notched=n

7\. gill-spacing: close=c,crowded=w,distant=d

8\. gill-size: broad=b,narrow=n

9\. gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g,
green=r,orange=o,pink=p,purple=u,red=e, white=w,yellow=y

10\. stalk-shape: enlarging=e,tapering=t

11\. stalk-root: bulbous=b,club=c,cup=u,equal=e,
rhizomorphs=z,rooted=r,missing=?

12\. stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s

13\. stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s

14\. stalk-color-above-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o,
pink=p,red=e,white=w,yellow=y

15\. stalk-color-below-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o,
pink=p,red=e,white=w,yellow=y

16\. veil-type: partial=p,universal=u

17\. veil-color: brown=n,orange=o,white=w,yellow=y

18\. ring-number: none=n,one=o,two=t

19\. ring-type: cobwebby=c,evanescent=e,flaring=f,large=l,
none=n,pendant=p,sheathing=s,zone=z

20\. spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r,
orange=o,purple=u,white=w,yellow=y

21\. population: abundant=a,clustered=c,numerous=n,
scattered=s,several=v,solitary=y

22\. habitat: grasses=g,leaves=l,meadows=m,paths=p,
urban=u,waste=w,woods=d

**Comprueba que todas las variables con de tipo cualitativo.**

```{r}
all(sapply(mushroom,is.factor))
```

# Train & Test

Utiliza la función `createDataPartition` de la librería `caret` para
separar el conjunto de datos en 80% de los datos para entrenamiento y
20% para test. Crea el vector `trainIndex` con los índices de los
registros para entreamiento

```{r}
library(caret, quietly = TRUE)
set.seed(666)
trainIndex <- createDataPartition(mushroom$class,p=0.8,times=1,list=F)

```

# Clasificador basado en reglas

## OneR

```{r}
library(OneR)
model_OneR <- OneR(class~.,data=mushroom[trainIndex,])
summary(model_OneR)
```

Evaluamos el modelo OneR sobre los conjuntos de entrenamiento y test

```{r}
class_hat_train <- predict(model_OneR,newdata=mushroom[trainIndex,])
class_hat_test <- predict(model_OneR,newdata=mushroom[-trainIndex,])
```

```{r}

confusionMatrix(class_hat_train,mushroom$class[trainIndex])
confusionMatrix(class_hat_test,mushroom$class[-trainIndex])
```

## CBA

Candidate classification association rules (CARs) are mined with the
standard APRIORI algorithm. Rules are ranked by confidence, support and
size. Then either the M1 or M2 algorithm are used to perform database
coverage pruning and to determin the number of rules to use and the
default class.

```{r}
library(arules)
library(arulesCBA)
#mushroomt<-as(mushroom,"transactions") #curiosidad
```

```{r}
model_CBA <- CBA(class~., mushroom[trainIndex,])
summary(model_CBA)
inspect(model_CBA$rules) 
```

Número de reglas que predicen cada clase

```{r}
inspect(subset(model_CBA$rules,subset=rhs %pin% c('e')))

inspect(subset(model_CBA$rules,subset=rhs %pin% c('p')))
```

Inspecciona las reglas con más Suporte y Confianza

```{r}
model_CBA <- CBA(class~., mushroom[trainIndex,],parameter = c(support=0.25,confidence=1),appearance='odor')
summary(model_CBA)
inspect(model_CBA$rules) 
```

Visualiza las reglas

```{r}
#rules.to.plot<-
library(arulesViz)
plot(model_CBA$rules,method='graph',engine='htmlwidget')
```

Evaluamos el modelo CBA sobre los conjuntos de entrenamiento y test

```{r}
class_hat_train <- predict(model_CBA,newdata=mushroom[trainIndex,])
class_hat_test <- predict(model_CBA,newdata=mushroom[-trainIndex,])

confusionMatrix(class_hat_train,mushroom$class[trainIndex])
confusionMatrix(class_hat_test,mushroom$class[-trainIndex])
```

