---
title: "Regla Única"
subitle: "Máster Oficial Ciencia Datos- UV"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
---

# Weather Play ¿Cúando jugamos al golf?

Cargamos el conjunto de datos *WeatherPlayData.xlsx* de la hoja "Categorical", fuerzamos la importacion a `data.frame` y observamos el contenido de su información.

```{r}
library(readxl)
weather.play<- read_excel("WeatherPlayData.xlsx",sheet = "Categorical")
weather.play<-as.data.frame(weather.play)
```

## ZeroR

ZeroR es el método de clasificación más sencillo que se basa en el objetivo e ignora todos los predictores. El clasificador ZeroR simplemente predice la categoría de mayoría (clase). Aunque no hay poder de predictibilidad en ZeroR, es útil para determinar el rendimiento base como punto de referencia para otros métodos de clasificación.

### Ejercicio

Determinamos el acierto aplicando el algoritmo ZeroR

```{r}
frq<-table(weather.play$Play)
max.frq <- which.max(frq)
model.zeror <- names(max.frq)
prediction<-rep(model.zeror,14)

table(prediction,weather.play$Play,useNA = "always",dnn=list("ZeroR","Play"))
mean(prediction==weather.play$Play)*100
acierto.ZR<-mean(prediction==weather.play$Play)

```

## OneR package

OneR es una buena idea para construir "el mejor modelo simple" que logre un equilibrio entre la mejor precisión posible con un modelo que sea lo suficientemente simple de entender.

<https://cran.r-project.org/web/packages/OneR/vignettes/OneR.html>

### Frecuencias

Determinamos cómo se distribuye la Clase para cada Predictor y buscamos las reglas que mejor predicen según el algoritmo **OneR**.

```{r}
xtabs(~Outlook+Play,data=weather.play)
```

### One R

Construir el modelo con el mejor predictor utilzando el paquete **OneR**.

```{r}
library(OneR)
model <- OneR(Play~.,data=weather.play, verbose = TRUE) #,ties.method="chisq")
```

Mostrar reglas aprendidas y diagnóstico de modelo.

```{r}
summary(model)
```

```{r}
library(vcd)
plot(model)
mosaicplot(~Outlook+Play,data=weather.play,shade=T)
```

Predecimos con el modelo y evaluamos el rendimiento.

```{r}
prediction<-predict(model,weather.play)
OneR::eval_model(prediction,actual=weather.play$Play)#
```

# Divorce Data Set 

<https://archive.ics.uci.edu/ml/datasets/Divorce+Predictors+data+set>

**Attribute Information:**

1\. If one of us apologizes when our discussion deteriorates, the discussion ends.
\
2. I know we can ignore our differences, even if things get hard sometimes.
\
3. When we need it, we can take our discussions with my spouse from the beginning and correct it.
\
4. When I discuss with my spouse, to contact him will eventually work.
\
5. The time I spent with my wife is special for us.
\
6. We don't have time at home as partners.
\
7. We are like two strangers who share the same environment at home rather than family.
\
8. I enjoy our holidays with my wife.
\
9. I enjoy traveling with my wife.
\
10. Most of our goals are common to my spouse.
\
11. I think that one day in the future, when I look back, I see that my spouse and I have been in harmony with each other.
\
12. My spouse and I have similar values in terms of personal freedom.
\
13. My spouse and I have similar sense of entertainment.
\
14. Most of our goals for people (children, friends, etc.) are the same.
\
15. Our dreams with my spouse are similar and harmonious.
\
16. We're compatible with my spouse about what love should be.
\
17. We share the same views about being happy in our life with my spouse
\
18. My spouse and I have similar ideas about how marriage should be
\
19. My spouse and I have similar ideas about how roles should be in marriage
\
20. My spouse and I have similar values in trust.
\
21. I know exactly what my wife likes.
\
22. I know how my spouse wants to be taken care of when she/he sick.
\
23. I know my spouse's favorite food.
\
24. I can tell you what kind of stress my spouse is facing in her/his life.
\
25. I have knowledge of my spouse's inner world.
\
26. I know my spouse's basic anxieties.
\
27. I know what my spouse's current sources of stress are.
\
28. I know my spouse's hopes and wishes.
\
29. I know my spouse very well.
\
30. I know my spouse's friends and their social relationships.
\
31. I feel aggressive when I argue with my spouse.
\
32. When discussing with my spouse, I usually use expressions such as \'you always\' or \'you never\' .
\
33. I can use negative statements about my spouse's personality during our discussions.
\
34. I can use offensive expressions during our discussions.
\
35. I can insult my spouse during our discussions.
\
36. I can be humiliating when we discussions.
\
37. My discussion with my spouse is not calm.
\
38. I hate my spouse's way of open a subject.
\
39. Our discussions often occur suddenly.
\
40. We're just starting a discussion before I know what's going on.
\
41. When I talk to my spouse about something, my calm suddenly breaks.
\
42. When I argue with my spouse, ı only go out and I don't say a word.
\
43. I mostly stay silent to calm the environment a little bit.
\
44. Sometimes I think it's good for me to leave home for a while.
\
45. I'd rather stay silent than discuss with my spouse.
\
46. Even if I'm right in the discussion, I stay silent to hurt my spouse.
\
47. When I discuss with my spouse, I stay silent because I am afraid of not being able to control my anger.
\
48. I feel right in our discussions.
\
49. I have nothing to do with what I've been accused of.
\
50. I'm not actually the one who's guilty about what I'm accused of.
\
51. I'm not the one who's wrong about problems at home.
\
52. I wouldn't hesitate to tell my spouse about her/his inadequacy.
\
53. When I discuss, I remind my spouse of her/his inadequacy.
\
54. I'm not afraid to tell my spouse about her/his incompetence.

`Class` debería ser un factor (aplica la lógica positiva!) y añade las etiquetas "yes" y "no". ¿importa el orden de los niveles del factor?

```{r}
library(readxl)
data <- read_excel("divorce.xlsx")
data<-as.data.frame(data)
table(data$Class)
#solucion
data$Class <- factor(data$Class,levels=c(0,1),labels=c('yes','no'))
```

## ZeroR

Utiliza la función `confusionMatrix` del paquete `caret` para determinar el rendimiento del modelo **ZeroR**. ¿cómo se denomina al acierto del modelo **ZeroR** entre las medidas que ofrece `confusionMatrix`?

```{r}
library(caret)
moda <- function(x) {
   return(names(which.max(table(x))))
}
#solucion


moda_div<-moda(data$Class) 

prediction<-rep(moda_div,dim(data)[1])

prediction_0 <- factor(prediction,levels=c('yes','no'))
confusionMatrix(prediction_0,data$Class)

```

## OneR

Discretiza los valores con la instrucción `bin` o `optbin` del paquete **OneR** para obtener el `data_disc`. ¿qué diferencia ambas funciones?

```{r}
library(OneR)
data_disc<-data
#solucion
data_disc <- optbin(Class~.,data=data_disc)
```

Utilizando la función **OneR** determina la regla que mejor estima el divorcio a partir de los datos discretizados `data_disc`.

```{r}
library(OneR)
#solucion
model <- OneR(Class~.,data=data_disc, verbose = TRUE)
summary(model)
```

Determina el rendimiento de la prueba con `confusionMatrix.` del paquete `caret`.

```{r}
# solucion
prediction <- predict(model,data_disc)
confusionMatrix(prediction,data$Class)
```
